Online Laterality Battery Pre-Processing - short version

### 2020-11-27 battery-preprocessing_v6_DB_split2.Rmd
revised manuscript

Modified to allow rerun on split half.
For this we take blocks of 4 rows, e.g. 1,2,3,4,9,10,11,12 etc.
This way the script will work for finger tapping as well as the other tasks.
Idea is to do this selection of rows right at the outset on the raw file.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
library(beeswarm) #just a kind of plot Dorothy likes
library(BSDA) #for ztest
```

@@

```{r demograhics, warning=FALSE}
#Reading from OSF - this only works if files are public
demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactors = F)

# relabel 
demo_dat$subject <- demo_dat$Participant.Public.ID
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject subject
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

@@

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- as.numeric( tmp[tmp$Question.Key == "Age",]$Response)
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

@@

allsum$genderpch <-2
w<-which(allsum$gender=='Male')
allsum$genderpch[w] <-1
w<-which(allsum$gender=='Prefer not to say')
allsum$genderpch[w]<-NA
```

@@

```{r sightednesss, warning=FALSE}
porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
miles_dat <-  read.csv("https://osf.io/ekqsp/download",stringsAsFactors = F)

miles_dat$test_subject <- "miles"
porta_dat$test_subject <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$Participant.Public.ID
sight_dat$subject <- as.factor(sight_dat$subject)

# count and print subjects
nsub <-length(levels(sight_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject subject
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

@@

    sightedness$subject[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and allsum
allsum <- merge(allsum, sightedness, by= "subject",all=TRUE)
head(allsum) #just to check all looking OK
```

@@

```{r EHI, warning=FALSE}
EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)

@@

# recode variables
EHI$subject <- EHI$Participant.Public.ID
EHI$subject <- as.factor(EHI$subject)

# reduce data
# new data
meaningful <- c("subject", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

@@

# re-reduce data
# new data
meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$subject))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("subject", "right_hand", "left_hand")

for (i in 1:nsub) { # loop through subjects
  myrow <- i
  subname <- levels(EHI$subject)[i] # find subject subject
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # sum for L hand
  }

@@

# now create a data frame only for index 
# merge with allsum
allsum$index_EHI <- NA #initialise
for (i in 1:nrow(allsum)){
  thissub <- allsum$subject[i]
  w<-which(EHI2$subject==thissub)
 allsum$index_EHI[i] <- EHI2$index_EHI[w]
}

@@

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(wsubjectth=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and subjectea to check EHI is valid etc.

@@

oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
allsum$handedness[oddcase] <- 'Left'
allsum$handpch[oddcase] <- 1
```

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)

# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$Participant.Public.ID
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")

for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject subject
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

@@

    lexTale_score$subject[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

@@

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale") + geom_vline(xintercept = 80, linetype="dashed", color = "black", size=1.5)
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

# now create a data frame only for index 
lexTALE <- dplyr::select(lexTale_score, "subject", "lexTALE")
# now merge with allsum
allsum <- merge(allsum, lexTALE, by= "subject",all=TRUE)
# now write the full allsum 
write.csv(allsum, file = "./allsum_out.csv")

@@

# mark those who ID don't pass the LexTALE
allsum <- allsum %>% mutate(lexOut= if_else(lexTALE < 80, "remove", "keep"))
```

@@

```{r readraw,include= FALSE, warning=FALSE}
# Set directory

  data_file <- "https://osf.io/mhk29/download" #day 1 data
  all_dat1 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)

    data_file <-  "https://osf.io/ebm6s/download"
     all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)

@@

alldat %>% filter(Attempt >= 1) # only answers 
# write as factors
alldat$subject <- as.factor(alldat$Participant.Public.ID)
alldat$RT <- as.numeric(as.character(alldat$Reaction.Time))
alldat$Trial <-as.numeric(alldat$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
alldat$Task.Name <- as.factor(alldat$Task.Name)
alldat$word<-substring(alldat$word,66,(nchar(alldat$word)-4)) #remove format from written word stimuli
oddeven <- 1 #set to zero for full dataset; to 1 for odds, and to 2 for evens
nr<-nrow(alldat)

if (oddeven==1){
  alldat<-alldat[alldat$Trial%%2==1,] #odd trials only
}
if (oddeven==2){
  alldat<-alldat[alldat$Trial%%2==0,] #even trials only
}

alldat$word_OG<-NA #create dummy column just to make it easier to align columns

@@

Create a simple numerical ID to make life easier when screening data if we need to identify subjects whose data is odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
We first rename the current subject variable so we can reuse this name with the numbered ID for all future computations.
Again, this is rather slow because done in a loop, which is probably not very efficient!

```{r makesubject}
#Use allsum as the basis for the check
#First rename all subject to origID
allsum <- allsum %>% rename(origID = subject) 
alldat <- alldat %>% rename(origID = subject)

alldat$subject <- NA

for (n in 1:nrow(allsum)){
  allsum$subject[n] <- n
  w<-which(alldat$origID == allsum$origID[n])
  alldat$subject[w] <-n

}
head(allsum)
```

@@

```{r excludelowCFacc}
#db modified to mark rather than drop cases with scores < .75
allsum$exclude <- 0 #default is to include

@@

print(paste('N flagged for exclusion because < 75% on at least one trial:',length(w)))
w <- union(which(allsum$Emot.corr.diff1 ==0),which(allsum$Emot.corr.diff2 == 0)) #also exclude if no correct response on one side
allsum$exclude[w]<-2 #code exclude = 2 if excluded because no response one side
```

@@

```{r dosummary}
#find included cases
w<-which(allsum$exclude==0)
inclsubs <- allsum$subject[w]

@@

Now remove outliers, using same function as for CF.
```{r dichotic.outliers}
dichotic_dat1 <- detect.outliers(dichotic_dat1,200,1.65) #use generic outlier removal function
outliertable1<-table(dichotic_dat1$subject,dichotic_dat1$outlier)
dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
outliertable2<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
#print(outliertable)
```

@@

#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
```

@@

```{r marktecherr}
techerr<-table(dichotic_dat2$subject)
te<-which(techerr<80)
techerr<-techerr[te] #subject numbers for those with few trials.
techerr<-as.numeric(names(techerr))
```
The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```
```{r add.dichotic.raw}
#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
DLdatL1<-dplyr::select(filter(DL_dat1,side=='Left'),subject,accurate)
DLdatR1<-dplyr::select(filter(DL_dat1,side=='Right'),subject,accurate)
DLdatL2<-dplyr::select(filter(DL_dat2,side=='Left'),subject,accurate)
DLdatR2<-dplyr::select(filter(DL_dat2,side=='Right'),subject,accurate)

@@

```{r rdt_outliers}
#First we create a RT column that is based on LOGRT
RDT_dat1$rawRT <- RDT_dat1$RT
RDT_dat2$rawRT <- RDT_dat2$RT
RDT_dat1$RT <- log(RDT_dat1$rawRT)
RDT_dat2$RT <- log(RDT_dat2$rawRT)
RDT_dat1<-detect.outliers(RDT_dat1,log(200),1.65) #numbers specify min RT and 
RDT_dat2<-detect.outliers(RDT_dat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
```

@@

```{r RDT_exclude}

w1x<-which(is.na(RDT_df1$RT))
w2x<-which(is.na(RDT_df2$RT))

@@

w1<-which(nr1$p.corr==0)
w2<-which(nr2$p.corr==0)

@@

w1a<-which(wr1$p.corr<75)
w2a<-which(wr2$p.corr<75)

@@

allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
allsum$exRDT[allsum$subject %in% exRDTx]<-2 #missing data
w<-which(allsum$exRDT==0)  
incRDT <- allsum$subject[w]
ninc <- length(incRDT)
print(paste0('RDT: number remaining after exclusions = ',ninc))

@@

```
```{r oldRDT_exclude}
#Original version now superseded
doold<-0
if(doold==1){
  wn <-unique(which(RDT_df1$accurate ==0),which(RDT_df2$accurate ==0)) #this catches those who were always wrong on 'no rhyme' trials, and possibly others where presentation messed up for technical reasons

w1<-which(RDT_df1$p.corr < 75)
w2<-which(RDT_df2$p.corr < 75)

exRDT<-sort(unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn])))

allsum$exRDT <- 0
allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
w<-which(allsum$exRDT==0)  
incRDT <- allsum$subject[w]
ninc <- length(incRDT)
print(paste0('RDT: number remaining after exclusions = ',ninc))

@@

# RT
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
}
```

@@

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
abline(0,1,lty=2)
text(-2,3.5,paste0('r = ',round(cor1,3)))
}
write.csv(allsum, paste0("LIs",oddeven,".csv"))
```

@@

>>>>

Modified to allow rerun on split half.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries

library(yarrr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
library(beeswarm) #just a kind of plot Dorothy likes
library(BSDA) #for ztest
```

@@

```{r demograhics, warning=FALSE}

oddeven <- 0 #set to zero for full dataset; to 1 for odds, and to 2 for evens
#NB we want to use exclusion criteria from full dataset for the split half as well, and so we need to run the script with oddeven set to zero before using it for split

#Demographics not needed for odd/even runs
if (oddeven==0){
#Reading from OSF - this only works if files are public
demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactors = F)

# relabel 
demo_dat$origID <- demo_dat$Participant.Public.ID
demo_dat$origID <- as.factor(demo_dat$origID)
# count and print subjects
nsub <-length(levels(demo_dat$origID))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("origID", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$origID)[i] # find origID subject
  myrows <- which(demo_dat$origID==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

@@

    allsum$origID[myrow] <- subname
    allsum$age[myrow] <- as.numeric( tmp[tmp$Question.Key == "Age",]$Response)
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

@@

allsum$genderpch <-2
w<-which(allsum$gender=='Male')
allsum$genderpch[w] <-1
w<-which(allsum$gender=='Prefer not to say')
allsum$genderpch[w]<-NA
}
```

@@

```{r sightednesss, warning=FALSE}
if (oddeven==0){
porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
miles_dat <-  read.csv("https://osf.io/ekqsp/download",stringsAsFactors = F)

miles_dat$test_origID <- "miles"
porta_dat$test_origID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$origID <- sight_dat$Participant.Public.ID
sight_dat$origID <- as.factor(sight_dat$origID)

# count and print subjects
nsub <-length(levels(sight_dat$origID))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("origID", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$origID)[i] # find origID subject
  myrows <- which(sight_dat$origID==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

@@

    sightedness$origID[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and allsum
allsum <- merge(allsum, sightedness, by= "origID",all=TRUE)
head(allsum) #just to check all looking OK
}
```

@@

```{r EHI, warning=FALSE}
if (oddeven==0){
EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)

@@

# recode variables
EHI$origID <- EHI$Participant.Public.ID
EHI$origID <- as.factor(EHI$origID)

# reduce data
# new data
meaningful <- c("origID", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

@@

# re-reduce data
# new data
meaningful <- c("origID", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$origID))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("origID", "right_hand", "left_hand")

for (i in 1:nsub) { # loop through subjects
  myrow <- i
  subname <- levels(EHI$origID)[i] # find origID subject
  myrows <- which(EHI$origID==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$origID[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # sum for L hand
  }

@@

# now create a data frame only for index 
# merge with allsum
allsum$index_EHI <- NA #initialise
for (i in 1:nrow(allsum)){
  thissub <- allsum$origID[i]
  w<-which(EHI2$origID==thissub)
 allsum$index_EHI[i] <- EHI2$index_EHI[w]
}

@@

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(width=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and ID to check EHI is valid etc.

@@

oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
allsum$handedness[oddcase] <- 'Left'
allsum$handpch[oddcase] <- 1
}
```

@@

```{r lexTALE, warning=FALSE}
if (oddeven==0){
# read data
lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)

# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$origID <- lexTALE$Participant.Public.ID
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$origID <- as.factor(lexTALE$origID)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("origID", "Word", "nonWord", "lexTALE")

for (i in 1:nsub) { # loop through origIDs
  subname <- levels(lexTALE$origID)[i] # find origID subject
  myrows <- which(lexTALE$origID==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

@@

    lexTale_score$origID[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

@@

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=origID, fill=origID)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale") + geom_vline(xintercept = 80, linetype="dashed", color = "black", size=1.5)
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

# now create a data frame only for index 
lexTALE <- dplyr::select(lexTale_score, "origID", "lexTALE")
# now merge with allsum
allsum <- merge(allsum, lexTALE, by= "origID",all=TRUE)
# now write the full allsum 
write.csv(allsum, file = "./allsum_out.csv")

@@

# mark those who ID don't pass the LexTALE
allsum <- allsum %>% mutate(lexOut= if_else(lexTALE < 80, "remove", "keep"))
}
```

@@

```{r readraw,include= FALSE, warning=FALSE}
  #reading direct from OSF is temperamental
  #Often gives error message if run from chunk, but line by line is OK. Grrr.
  readosf<-0
  if(readosf==1){
  data_file <- "https://osf.io/mhk29/download" #day 1 data
  all_dat1 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)

    data_file <-  "https://osf.io/ebm6s/download"
     all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
  }
  if(readosf==0){
    all_dat1=read.csv("./data/day1_combined.csv",na.strings = "NA",stringsAsFactors=F)
      all_dat2=read.csv("./data/day2_combined.csv",na.strings = "NA",stringsAsFactors=F)
  }

@@

alldat %>% filter(Attempt >= 1) # only answers

# write as factors
alldat$origID <- as.factor(alldat$Participant.Public.ID)
alldat$subject<-as.numeric(alldat$origID)
alldat$RT <- as.numeric(as.character(alldat$Reaction.Time))
alldat$Trial <-as.numeric(alldat$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
alldat$Task.Name <- as.factor(alldat$Task.Name)
alldat$word<-substring(alldat$word,66,(nchar(alldat$word)-4)) #remove format from written word stimuli

##################################################################################
#      MODIFICATION FOR ODD AND EVEN ANALYSIS
##################################################################################

nr<-nrow(alldat)

if (oddeven==1){
  alldat<-alldat[alldat$Trial%%2==1,] #odd trials only
}
if (oddeven==2){
  alldat<-alldat[alldat$Trial%%2==0,] #even trials only
}
```

```{r reuse.allsum}
#for odd and even subsets, read in original allsum and select demographic and exclusion cols to keep
if (oddeven>0){
  allsum<-read.csv("./data/allsum_out.csv",stringsAsFactors=F)
  keepcols<-c("subject" , "origID","age" ,"gender","footedness" ,"handedness","bilingual", "strongest.language","handpch","genderpch" ,"miles", "porta","index_EHI" ,"lexTALE", "lexOut","exclude","exRDT"      )
  allsum<-allsum[,keepcols]
}
head(allsum)

```
```{r subnums}
##################################################################################
#find subject number corresponding to orig ID

allsum$subject<-NA
for (i in 1:nrow(allsum)){
  thisID<-allsum$origID[i]
  w<-which(alldat$origID == thisID)
  allsum$subject[i]<-alldat$subject[w[1]]
}

@@

@@

```{r excludelowCFacc}
if (oddeven==0) {
#db modified to mark rather than drop cases with scores < .75
allsum$exclude <- 0 #default is to include

@@

print(paste('N flagged for exclusion because < 75% on at least one trial:',length(w)))
w <- union(which(allsum$Emot.corr.diff1 ==0),which(allsum$Emot.corr.diff2 == 0)) #also exclude if no correct response on one side
allsum$exclude[w]<-2 #code exclude = 2 if excluded because no response one side
}
```

@@

```{r dosummary}
#find included cases
w<-which(allsum$exclude==0)

inclsubs <- allsum$subject[w]

@@

Now remove outliers, using same function as for CF.
```{r dichotic.outliers}
if (oddeven==0){
dichotic_dat1 <- detect.outliers(dichotic_dat1,200,1.65) #use generic outlier removal function
outliertable1<-table(dichotic_dat1$subject,dichotic_dat1$outlier)
dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
outliertable2<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
#print(outliertable)
}
```

@@

if(oddeven==0){
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
}
```

@@

```{r marktecherr}
Nfortech<-80 #need at least this N trials
if(oddeven>0){Nfortech<-40} #half the number for split half runs
techerr<-table(dichotic_dat2$subject)
te<-which(techerr<Nfortech)
techerr<-techerr[te] #subject numbers for those with few trials.
techerr<-as.numeric(names(techerr))
w<-which(allsum$subject%in% techerr)
allsum$exclude[w]<-4 #exclude for tech error

```
The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
#only plot/t-test included cases
origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```
```{r add.dichotic.raw}
#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
DLdatL1<-dplyr::select(filter(DL_dat1,side=='Left'),subject,accurate)
DLdatR1<-dplyr::select(filter(DL_dat1,side=='Right'),subject,accurate)
DLdatL2<-dplyr::select(filter(DL_dat2,side=='Left'),subject,accurate)
DLdatR2<-dplyr::select(filter(DL_dat2,side=='Right'),subject,accurate)

@@

```{r rdt_outliers}

#First we create a RT column that is based on LOGRT
RDT_dat1$rawRT <- RDT_dat1$RT
RDT_dat2$rawRT <- RDT_dat2$RT
RDT_dat1$RT <- log(RDT_dat1$rawRT)
RDT_dat2$RT <- log(RDT_dat2$rawRT)
if(oddeven==0){
RDT_dat1<-detect.outliers(RDT_dat1,log(200),1.65) #numbers specify min RT and 
RDT_dat2<-detect.outliers(RDT_dat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
}
#for splithalves read in the original RDT file for all trials and then split
#This means we keep the outlier designation as before
if(oddeven>0){
  tempRDT1<-read.csv("./data/myRDTdat1.csv",stringsAsFactors=F)
  tempRDT2<-read.csv("./data/myRDTdat2.csv",stringsAsFactors=F)
  if(oddeven==1){
  RDT_dat1<-tempRDT1[tempRDT1$Trial%%2==1,] #odd trials only
  RDT_dat2<-tempRDT2[tempRDT2$Trial%%2==1,] #odd trials only
  }
    if(oddeven==2){
  RDT_dat1<-tempRDT1[tempRDT1$Trial%%2==0,] #even trials only
  RDT_dat2<-tempRDT2[tempRDT2$Trial%%2==0,] #even trials only
  }
}
```

@@

```{r RDT_exclude}
if (oddeven==0){
w1x<-which(is.na(RDT_df1$RT))
w2x<-which(is.na(RDT_df2$RT))

@@

w1<-which(nr1$p.corr<75)
w2<-which(nr2$p.corr<75)

@@

w1a<-which(wr1$p.corr<50)
w2a<-which(wr2$p.corr<50)

@@

allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
allsum$exRDT[allsum$subject %in% exRDTx]<-2 #missing data
}

w<-which(allsum$exRDT==0)

incRDT <- allsum$subject[w]
ninc <- length(incRDT)
print(paste0('RDT: number remaining after exclusions = ',ninc))

@@

@@

# RT
origdat$X1 <- origdat$RT
measure <- 'RT'
#need to exclude those who don't have RT values on both L and R
w<-which(is.na(origdat$X1))
if (length(w)>0){
  takeout<-unique(origdat$subject[w])
  ww<-which(origdat$subject %in% takeout)
  origdat<-origdat[-ww,]
}
dopirate(origdat,task,measure)
}
```

@@

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
abline(0,1,lty=2)
text(-2,3.5,paste0('r = ',round(cor1,3)))
}
write.csv(allsum, paste0("./data/LIs",oddeven,".csv"))
```

@@

```{r finalbit}
if(oddeven==0){
